---
title: "Breastcancer Dataset BIOINFO"
author: "Ha Huy Hoang and Nguyen Anh Thu"
date: "4/23/2020"
output:
html_document:
  code_download: yes
  df_print: kable
  highlight: tango
---

# Introduction

This is one of three domains provided by the Oncology Institute that has repeatedly appeared in the machine learning literature.

This data set includes 201 instances of one class and 85 instances of another class. The instances are described by 9 attributes, some of which are linear and some are nominal.

# Library
```{r message=FALSE, warning=FALSE}
library(dplyr)
library(car)
library(MASS)
library(ggplot2)
library(Hmisc)
library(gridExtra)
library(psych)
```

# Data
```{r}
data = read.csv("https://raw.githubusercontent.com/pnhuy/bioinfo/master/datasets/breast_cancer/breast-cancer.csv")
head(data)
```

```{r}
str(data)
```

```{r}
data=na.omit(data)
summary(data)
```
```{r}
describe(data)
```

## Class vs age
```{r}
table(data$Class,data$age)
```
```{r}
pairs.panels(data)
```

# Plot
```{r}
data = filter(data, breast.quad != '?' & node.caps != '?')
```

## Bar plots for one variable
```{r fig.height=12, fig.width=18}
p1 = ggplot(data, aes(age)) + geom_bar(aes(fill=age))+labs(title = "Age")
p2 = ggplot(data, aes(Class)) +geom_bar(aes(fill=Class))+labs(title = "Class")
p3 = ggplot(data, aes(irradiat)) + geom_bar(aes(fill=irradiat))+labs(title = "irradiat")
p4 = ggplot(data, aes(breast.quad)) + geom_bar(aes(fill=breast.quad))+labs(title = "breast.quad")
p5 = ggplot(data, aes(breast)) + geom_bar(aes(fill=breast))+labs(title = "breast")
p6 = ggplot(data, aes(node.caps)) + geom_bar(aes(fill=node.caps))+ labs(title = "node.caps")
p7 = ggplot(data, aes(menopause)) + geom_bar(aes(fill=menopause))+labs(title = "menopause")
p8 = ggplot(data, aes(tumor.size)) + geom_bar(aes(fill=tumor.size))+labs(title = "Tumor Size")
p9 = ggplot(data, aes(inv.nodes)) + geom_bar(aes(fill=inv.nodes))+labs(title = "inv nodes")
grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8,p9)
```

## Plot of multiple variables
```{r}
ggplot(data=data,aes(x=age,fill=irradiat,))+geom_bar()
ggplot(data,aes(tumor.size))+geom_bar()+facet_wrap(~irradiat)+theme(axis.text.x = element_text(angle = 90))+ggtitle("Irradiation Based on Tumor Size")+xlab("Tumor Size")+ylab("Count")
ggplot(data,aes(x=Class,fill=irradiat))+geom_bar(position ="dodge")
ggplot(data,aes(x=tumor.size,fill=menopause))+geom_bar(position="dodge")
ggplot(data,aes(x=age,fill=tumor.size))+geom_bar(position ="dodge")
```

## Piechart
```{r}
ggplot(data,aes(x="",fill=menopause)) + geom_bar(width=3)+coord_polar("y")
```

# Density plot
```{r}
ggplot(data,aes(x=`deg.malig`,fill=`tumor.size`))+geom_density(alpha=0.4)+ggtitle("Deg Malignant vs Tumor size")+xlab("Deg Malignant")+ylab("Density")
```


## Labeling features
```{r}
sum(is.na(data))

#Fix NA in nodecaps
data$node.caps=ifelse(is.na(data$node.caps),ave(data$node.caps,FUN=function(x)"no"),data$node.caps)
#Fix NA in breasquad
data$breast.quad=ifelse(is.na(data$breast.quad),ave(data$breast.quad,FUN=function(x)"left_low"),data$breast.quad)
#Check levels
factor (data$irradiat)
#labeling 
data$Class= factor (data$Class, labels= c(0,1) , levels= c("no-recurrence-events", "recurrence-events"))

data$age= factor (data$age,labels= c(0,1,2,3,4,5) , levels= c("20-29","30-39","40-49","50-59","60-69","70-79")) 

data$menopause= factor (data$menopause, labels= c(0,1,2) , levels= c("premeno","ge40","lt40")) 

data$tumor.size= factor (data$tumor.size, labels= c(0,1,2,3,4,5,6,7,8,9,10) , levels= c("0-4","5-9","10-14", "15-19", "20-24","25-29", "30-34", "35-39","40-44","45-49","50-54")) 

data$inv.nodes= factor (data$inv.nodes, labels= c(0,1,2,3,4,5,6) , levels= c("0-2","3-5","6-8", "9-11","12-14", "15-17", "24-26"))

data$node.caps= factor (data$node.caps, labels= c(0,1) , levels= c("2", "3"))

data$deg.malig= factor (data$deg.malig, labels= c(0,1,2) , levels= c("1", "2", "3"))


data$breast= factor (data$breast, labels= c(0,1) , levels= c("left", "right")) 

data$breast.quad= factor (data$breast.quad, labels= c(0,1,2,3,4) , levels= c("2", "3", "4", "5", "6")) 

data$irradiat= factor (data$irradiat, labels= c(0,1) , levels= c("no", "yes")) 
```


# Testing
## Ttest
```{r}
#Testing tumorsize vs irradiat
t.test(as.numeric(data$tumor.size)~data$irradiat)
```

```{r}
#Test tumorsize vs nodecaps
t.test(as.numeric(data$tumor.size)~data$node.caps)
```

## Anova
```{r}
ANOVA = aov(as.numeric(data$tumor.size)~data$node.caps, data=data)
summary(ANOVA)
TukeyHSD(ANOVA)
```

# Model
```{r}
model <- lm(as.numeric(data$tumor.size) ~ ., data=data)
step.model <- stepAIC(model, direction = "both", 
                      trace = FALSE)
summary(step.model)
```
# Interpret the model
## Variables names
+ node.caps1,breast.quad1,breast.quad2,breast.quad4 are $X_{1}$,$X_{2}$,$X_{3}$,$X_{4}$, respectively.
+ tumor.size is Y.
The model above is achieved by using the lm() function in R and the output is called using the summary() function on the model.
Formula Call:
  $Y=4,14+0,69X_{1} +1,36X_{2} +1,56X_{3} +2,20X_{4}$
  
Residuals : break down into 5 sumary point:  Min, Max, Median, 1Q,3Q. In the model, we can see that the distribution of the residuals do not appear to be strongly symetrical. 

Cofficients : reprensent the intercept, slope terms, P-value
Another results : R-square = 0.1368, Adjusted R squared = 0.1144, F statistic = 6.092

## Predict
The following section, we will calculate Y hat:
```{r}
head(predict(step.model),10)
```
Finally, we will have Residual testing:
```{r fig.height=6, fig.width=8}
op = par(mfrow = c(2,2))
p12 = plot(step.model, c(1,2))
p13 = spreadLevelPlot(step.model)
```

As you can see on three images above:
  
+ The mean of residuals is approximately zero
+ Variance is always constant
+ Follow a normal distribution

